---
title: "Oregon Campaign Response on Twitter"
author: "Hannah Metzler"
date: "17 Nov 2020"
output: 
  pdf_document:
    toc_depth: 2
    df_print: kable
    keep_tex: true
url_colour: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

# session settings
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
options(scipen=10)
```

```{r}
#packages
library(dplyr)
library(ggplot2) #for figures
library(scales)
library(cowplot) #for figures, placing next to each other
```

```{r,data}
# Run data preprocessing the first time you use this script (takes about 1 minute or so)
# source("scripts/01preparedata.R")

# or load data instead once data has been preprocessed
df <- read.csv("data/TweetVolumeLongOfEffectExcludingCampaignAuthors.csv") %>% 
  mutate(date=as.Date(date), 
         state=factor(state),
         period = factor(period, levels = c("Baseline","2019 before campaign", "Campaign-week", "After"), ordered=TRUE))

```

# Keywords used to search tweets in the query for Crimson Hexagon

We used two different queries:

* Label "Campaign": language:en AND region:USA.Or AND suicid* AND (breakingthesilenceor.com OR "#BreakingTheSilenceOR" OR "@BreakSilenceOR" OR "Breaking the silence" OR #Breakingthesilence) 
* Label "Prevention": language:en AND region:USA.Or AND suicid* AND (800273talk OR 18002738255 OR "1-800-273-TALK" OR "1-800-273" OR "1-800-273-8255" OR lifeline OR "suicide hotline" OR "suicide prevention hotline" OR Betheoneto OR #BeThe1To OR #Betheoneto OR Bethe1to) 

**Additional query settings**

* We use each query twice, once while excluding the campaign initiators as authors of tweets, and once including them. To exclude, we added "AND NOT (author:breakSilenceOr OR author:800273TALK OR author:Lines_for_Life)". 
* To download only English tweets from Oregon or Washington, we additionally used the query specifications: language:en AND region:USA.Or (or instead USA.WA). 


```{r, time series function}
plotTS <- function(df, mode="percent_change", title="Country TS", limits=NULL, factor=15, textsize=13, axistextsize=12)
{
  
  #make plot
  plt <- ggplot(data=df)
  
  if (mode == "percent_change")
    plt <- plt +
      geom_ribbon(aes(x=date, ymin=(prlow-bl)/bl, ymax=(prhigh-bl)/bl, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line( aes(x=date, y=(pr-bl)/bl, colour=keywords))+
      ylab("% Difference over baseline")
  if (mode == "percent_abs")
    plt <- plt +
      geom_ribbon(aes(x=date, ymin=prlow, ymax=prhigh, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line( aes(x=date, y=pr, colour=keywords), alpha=0.4)+
      ylab("% tweets with keywords") 
  if (mode == "Z")
    plt <- plt +
      geom_ribbon(aes(x=date, y=(pr-bl)/sd, ymin=(prlow-bl)/sd, ymax=(prhigh-bl)/sd, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line(aes(x=date, y=(pr-bl)/sd, colour=keywords))+
      ylab("Z-score") #y axis label
  
  plt <- plt +  ggtitle(title)+ #plot title
    scale_fill_manual(  values = c("Campaign wo authors"="blue", "Campaign with authors"="cyan","Prevention wo authors"="red", "Prevention with authors"="orange", alpha=0.4), name="Keywords in tweets")+
    scale_colour_manual(  values = c("Campaign wo authors"="blue", "Campaign with authors"="cyan","Prevention wo authors"="red", "Prevention with authors"="orange"), name="Keywords in tweets")+
    theme_bw()+ theme(text=element_text(size=textsize), axis.text=element_text(size=axistextsize), 
                      axis.title.x = element_blank(),
                      axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1), 
                      legend.position=c(0.75,0.7),
                      plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"), 
                      panel.grid.minor= element_blank())+
    scale_x_date(date_breaks="1 month", date_labels = "%b")+
    geom_hline(yintercept = 0, colour = "grey50")+ #mean line
    geom_vline(xintercept = as.Date("2019-04-05"), colour = "grey50", linetype=2)+ #campaign week start
    geom_vline(xintercept = as.Date("2019-04-14"), colour = "grey50", linetype=2) #campaign week end
  
  if (!is.null(limits))
    plt <- plt + 
    scale_y_continuous(limits = limits)
  
  
  return(plt)
}

```

\newpage



# Time series around campaign week: absolute percentages of tweets with at least one keyword

```{r, ts absolute campaign, fig.height=8, fig.width=6}
df_or <- df %>% 
  filter(state=="Oregon" & date>"2019-03-01"&date<"2019-06-01") 
ts_or <- plotTS(df=df_or, mode="percent_abs", title="Oregon")+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

df_wa <- df %>% 
  filter(state=="Washington" & date>"2019-03-01"&date<"2019-06-01")
ts_wa <- plotTS(df=df_wa,  mode="percent_abs", title="Washington")+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

plot_grid(ts_or, ts_wa, ncol=1)
```

* Only a few tweets were posted from accounts of the campaign initiators (BreakSilenceOR, Lines_for_Life, 800273TALK) themselves. Most of the social media response in Oregon remains if these are excluded as authors. 
* For tweets aimed at prevention (including lifeline numbers, terms, or the terms/hashtags related to "BeTheOneToo"), the difference in percentage is so minimal that it is not visible. There are only `r sum(with(df, subset(pr, keywords == "Prevention wo authors")) != with(df, subset(pr, keywords == "Prevention with authors")))` tweets more from these authors distributed across the entire time period and both states.
* This means we can exclude tweets form these three twitter accounts, the response on Twitter is not only explained by them. 


\newpage

# Do Campaign tweets occur at all during the baseline?

Limit the y-scale to check if small percentages below 0.05% occur: 

```{r, campaign tweets below 0.05 before campaign, fig.height=8, fig.width=7}
fontsingleplots = 10
ggplot(data=filter(df, keywords=="Campaign wo authors" | keywords=="Campaign with authors"))+
  geom_line( aes(x=date, y=pr, colour=keywords))+
  facet_grid(rows=vars(state))+
  ylab("% tweets with keywords") +
  ggtitle("Entire period")+ #plot title
  scale_fill_manual(  values = c("Campaign wo authors"="blue", "Campaign with authors"="cyan","Prevention wo authors"="red", "Prevention with authors"="orange", alpha=0.4), name="Keywords in tweets")+
  scale_colour_manual(  values = c("Campaign wo authors"="blue", "Campaign with authors"="cyan","Prevention wo authors"="red", "Prevention with authors"="orange"), name="Keywords in tweets")+
  theme_bw()+ theme(text=element_text(size=fontsingleplots), axis.text=element_text(size=fontsingleplots), 
                    axis.title.x = element_blank(),
                    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1), 
                    legend.position=c(0.2,0.9),
                    plot.margin = margin(0.5, 0.5, 2, 0.5, "cm"), 
                    panel.grid.minor= element_blank())+
  scale_x_date(date_breaks="1 month", date_labels = "%b")+
  geom_hline(yintercept = 0, colour = "grey50")+ #mean line
  geom_vline(xintercept = as.Date("2019-04-05"), colour = "grey50", linetype=2)+ #campaign week start
  geom_vline(xintercept = as.Date("2019-04-14"), colour = "grey50", linetype=2)+ #campaign week end
  ylim(0, 0.045)
```

\newpage

To look at the small frequencies before and after the campaign, I cut the y-scale to 0.002 percent max (1.0 would be 1%). Because of this, the high frequencies during the campaign week are not displayed, and the campaign week window is empty. This shows that breaking the silence terms were used on 3 or 4 days (depending on the state) before the campaign, but at a negligible percentage below 0.0005%. 

```{r, campaign tweets below 0.002 before campaign, fig.height=8, fig.width=7}
ggplot(data=filter(df, keywords=="Campaign wo authors" | keywords=="Campaign with authors"))+
  geom_line( aes(x=date, y=pr, colour=keywords))+
  facet_grid(rows=vars(state))+
  ylab("% tweets with keywords") +
  ggtitle("Entire period, y-scale cut ")+ #plot title
  scale_colour_manual(  values = c("Campaign wo authors"="blue", "Campaign with authors"="cyan","Prevention wo authors"="red", "Prevention with authors"="orange"), name="Keywords in tweets")+
  theme_bw()+ theme(text=element_text(size=fontsingleplots), axis.text=element_text(size=fontsingleplots), 
                    axis.title.x = element_blank(),
                    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1), 
                    legend.position=c(0.2,0.9),
                    plot.margin = margin(0.5, 0.5, 2, 0.5, "cm"), 
                    panel.grid.minor= element_blank())+
  scale_x_date(date_breaks="2 month", date_labels = "%b")+
  geom_hline(yintercept = 0, colour = "grey50")+ #mean line
  geom_vline(xintercept = as.Date("2019-04-05"), colour = "grey50", linetype=2)+ #campaign week start
  geom_vline(xintercept = as.Date("2019-04-14"), colour = "grey50", linetype=2)+ #campaign week end
  ylim(0, 0.002)
```

\newpage

# Descriptive statistics

## Total sample

```{r}
df %>% 
  group_by(keywords, state) %>% 
  summarize(meanpr = mean(pr), sdpr = sd(pr), meanbl = mean(bl), meancilow = mean(prlow), meancihigh = mean(prhigh))
```

## Sample size and mean per day per state

```{r}
df %>% 
  group_by(state) %>%
  filter(keywords=="Campaign wo authors") %>% #choose only one of the repeated datasets (tot long the same every time)
  summarize(milliontweets = round(sum(as.numeric(tot_n)/1000000), 2),
            meantweetsperday = round(mean(as.numeric(tot_n)), 2)) %>% 
  ungroup()
```


## Number of tweets per query per time period in Oregon

* Baseline: 24 September 2018 to 24 March 2019 (half a year ending 2 weeks before the campaign)
* Campaign-week: 7-14 April 2019
* After: 15 April until 15 October 2019 (half a year after the campaign)

```{r}
df %>% 
  filter(state=="Oregon") %>% 
  group_by(period, keywords) %>%
  summarize(meantweetsperday = round(mean(as.numeric(keywords_n)), 2),
            sd=round(sd(keywords_n), 2)) %>% 
  ungroup()
```

## Number of tweets per query per time period in Washington

```{r}
df %>% 
  filter(state=="Washington") %>% 
  group_by(period, keywords) %>%
  summarize(meantweetsperday = round(mean(as.numeric(keywords_n)), 2),
            sd=round(sd(keywords_n), 2)) %>% 
  ungroup()
```
