---
title: "Oregon Campaign Response on Twitter"
author: "Hannah Metzler"
date: "11/9/2020"
output: 
  pdf_document:
    toc_depth: 2
    df_print: kable
    keep_tex: true
url_colour: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

# session settings
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
options(scipen=10)
```

```{r}
#packages
library(dplyr)
library(ggplot2) #for figures
library(scales)
# library(gridExtra)
library(cowplot) #for figures, placing next to each other
```

```{r,data}
# Run data preprocessing the first time you use this script (takes about 1 minute or so)
# source("scripts/01preparedata.R")

# or load data instead once data has been preprocessed
df <- read.csv("data/TweetVolumeLong.csv") %>% 
  mutate(date=as.Date(date), 
         state=factor(state), 
         period = factor(period, levels = c("Baseline", "Campaign-week", "After")))
```

# Keywords used to search tweets in the query for Crimson Hexagon

The keywords associated with the campaign were: breakingthesilenceor.com,  "#BreakingTheSilenceOR",  @BreakSilenceOR (the Twitter account of the campaign),  800273talk,  18002738255,  "1-800-273-TALK",  "1-800-273",  "1-800-273-8255" (all lifeline numbers),  "Breaking the silence",  #Breakingthesilence,  Betheoneto,  #BeThe1To,  #Betheoneto,  Bethe1to,  lifeline,  “suicide hotline”. We used these keywords either togetheter with or without the terms "suicide" and "suicidal": 

* Label "Campaign": only these keywords
* Label "Campaign + Suicide": these keywords AND suicid*

To download only English tweets from Oregon or Washington, we additionally used the query specifications: language:en AND region:USA.Or (or instead USA.WA). 


```{r, time series function}
plotTS <- function(df, mode="percent_change", title="Country TS", limits=NULL, factor=15, textsize=13, axistextsize=12)
{
  
  #make plot
  plt <- ggplot(data=df)
  
  if (mode == "percent_change")
    plt <- plt +
      geom_ribbon(aes(x=date, ymin=(prlow-bl)/bl, ymax=(prhigh-bl)/bl, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line( aes(x=date, y=(pr-bl)/bl, colour=keywords))+
      ylab("% Difference over baseline")
  if (mode == "percent_abs")
    plt <- plt +
      geom_ribbon(aes(x=date, ymin=prlow, ymax=prhigh, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line( aes(x=date, y=pr, colour=keywords))+
      ylab("% tweets with keywords") 
  if (mode == "Z")
    plt <- plt +
      geom_ribbon(aes(x=date, y=(pr-bl)/sd, ymin=(prlow-bl)/sd, ymax=(prhigh-bl)/sd, fill=keywords, colour=keywords), alpha=0.2, size=0) +
      geom_line(aes(x=date, y=(pr-bl)/sd, colour=keywords))+
      ylab("Z-score") #y axis label
  
  plt <- plt +  ggtitle(title)+ #plot title
    scale_fill_manual(values = c("Campaign"="orange", "Campaign + Suicide"="cyan"), name="Keywords in tweets")+
    scale_colour_manual(values = c("Campaign"="orange", "Campaign + Suicide"="cyan"), name="Keywords in tweets")+
    theme_bw()+ theme(text=element_text(size=textsize), axis.text=element_text(size=axistextsize), 
                      axis.title.x = element_blank(),
                      axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1), 
                      legend.position=c(0.7,0.77),
                      plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"), 
                      panel.grid.minor= element_blank())+
    scale_x_date(date_breaks="1 month", date_labels = "%b-%y")+
    geom_hline(yintercept = 0, colour = "grey50")+ #mean line
    geom_vline(xintercept = as.Date("2019-04-05"), colour = "grey50", linetype=2)+ #campaign week start
    geom_vline(xintercept = as.Date("2019-04-14"), colour = "grey50", linetype=2) #campaign week end
  
    if (!is.null(limits))
    plt <- plt + 
    scale_y_continuous(limits = limits)
  
  
  return(plt)
}

```

\newpage

# Time series around campaign week: percent change

Peaks coincide with these events (suicide cases in red and suicide prevention events in green)

* March 23 and 25 2019: peak before the campaign: Two Parkland survivors, and then the father of a Sandy Hook victim die by suicide ([source](https://www.ctvnews.ca/world/two-parkland-survivors-father-of-sandy-hook-victim-die-by-suicide-1.4351198))

```{r, ts change campaign, fig.height=8, fig.width=6.5}
df_or <- df %>% 
  filter(state=="Oregon" & date>"2019-03-01"&date<"2019-06-01")
ts_or <- plotTS(df=df_or, title="Oregon", limits=c(-1,49))+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

df_wa <- df %>% 
  filter(state=="Washington" & date>"2019-03-01"&date<"2019-06-01")
ts_wa <- plotTS(df=df_wa, title="Washington", limits=c(-1,49))+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

plot_grid(ts_or, ts_wa, ncol=1)
```
 * The change in percentage of tweets with keywords compared to baseline is much larger when we look at tweets that contain both the campaign words and suicid*. This shows that most tweets referring to the campaign mentioned suicide (e.g. not just saying call the lifeline, but saying call the suicide lifeline). 
 * the change is very clear in Oregon, and entirely absent in Washington
\newpage

# Time series around campaign week: absolute percentages of tweets with at least one keyword

```{r, ts absolute campaign, fig.height=8, fig.width=6.5}
df_or <- df %>% 
  filter(state=="Oregon" & date>"2019-03-01"&date<"2019-06-01")
ts_or <- plotTS(df=df_or, mode="percent_abs", title="Oregon")+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

df_wa <- df %>% 
  filter(state=="Washington" & date>"2019-03-01"&date<"2019-06-01")
ts_wa <- plotTS(df=df_wa,  mode="percent_abs", title="Washington")+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=2)

plot_grid(ts_or, ts_wa, ncol=1)
```

\newpage

# Time series one year before the campaign: Percent change

1 year before and half a year after the campaign

Peaks coincide with these events (suicide cases in red and suicide prevention events in green):

* March 23 and 25 2019: peak before the campaign: Two Parkland survivors, and then the father of a Sandy Hook victim die by suicide ([source](https://www.ctvnews.ca/world/two-parkland-survivors-father-of-sandy-hook-victim-die-by-suicide-1.4351198))
* June 5 2018: suicide of Kate Spade
* June 8 2018: suicide of Anthony Bourdain
* September 10 2018/2019: world suicide prevention day
* December 16 2018: 3-digit suicide hotline number gets introduced
* February 19 2019: transgender lifeline gets introduced


```{r, ts entire period, fig.height=15, fig.width=14}
df_or <- df %>% 
  filter(state=="Oregon" & date < "2019-05-01")
ts_or <- plotTS(df=df_or, title="Oregon - 1 year before", limits=c(-1,49), textsize=20, axistextsize = 16)+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=4)+# suicides school shooting victims and bereaved
  geom_vline(xintercept = as.Date("2018-06-05"), colour = "red3", linetype=4)+# Kate Spade
   geom_vline(xintercept = as.Date("2018-06-08"), colour = "red3", linetype=4)+ # Bourdain
  geom_vline(xintercept = as.Date("2018-09-10"), colour = "green2", linetype=4)+# World suicide prevention day
 geom_vline(xintercept = as.Date("2018-12-16"), colour = "green2", linetype=4)+# 3-digit suicide hotline number gets introduced
  geom_vline(xintercept = as.Date("2019-02-18"), colour = "green2", linetype=4)+ #transgender lifeline gets introduced
  geom_vline(xintercept = as.Date("2019-09-10"), colour = "green2", linetype=4)# World suicide prevention day

df_wa <- df %>% 
  filter(state=="Washington" & date < "2019-05-01")
ts_wa <- plotTS(df=df_wa, title="Washington - 1 year before", limits=c(-1,49), textsize=20, axistextsize = 16)+
  geom_vline(xintercept = as.Date("2019-03-24"), colour = "red3", linetype=4)+ # suicides school shooting victims and bereaved
   geom_vline(xintercept = as.Date("2018-06-05"), colour = "red3", linetype=4)+# Kate Spade
  geom_vline(xintercept = as.Date("2018-06-08"), colour = "red3", linetype=4)+ # Bourdain
   geom_vline(xintercept = as.Date("2018-09-10"), colour = "green2", linetype=4)+# World suicide prevention day
  geom_vline(xintercept = as.Date("2018-12-16"), colour = "green2", linetype=4)+# 3-digit suicide hotline number gets introduced
  geom_vline(xintercept = as.Date("2019-02-18"), colour = "green2", linetype=4)+ #transgender lifeline gets introduced
  geom_vline(xintercept = as.Date("2019-09-10"), colour = "green2", linetype=4)# World suicide prevention day

plot_grid(ts_or, ts_wa, ncol=1)
```

# Time series after the campaign: Percent change

* Green line: world suicide prevention day

```{r, ts after campaign, fig.height=15, fig.width=14}
df_or <- df %>% 
    filter(state=="Oregon" & date>"2019-04-01")
ts_or <- plotTS(df=df_or, title="Oregon - After campaign", limits=c(-1,49), textsize=25, axistextsize = 16)+
  geom_vline(xintercept = as.Date("2019-09-10"), colour = "green2", linetype=4)# World suicide prevention day

df_wa <- df %>% 
  filter(state=="Washington"& date>"2019-04-01")
ts_wa <- plotTS(df=df_wa, title="Washington - After campaign", limits=c(-1,49), textsize=25, axistextsize = 16)+
  geom_vline(xintercept = as.Date("2019-09-10"), colour = "green2", linetype=4)# World suicide prevention day

plot_grid(ts_or, ts_wa, ncol=1)
```

\newpage



# Descriptive statistics

## Total sample

```{r}
df %>% 
  group_by(keywords, state) %>% 
  summarize(meanpr = mean(pr), sdpr = sd(pr), meanbl = mean(bl), meancilow = mean(prlow), meancihigh = mean(prhigh))
```

## Sample size and mean per day per state

```{r}
df %>% 
  group_by(state) %>%
  filter(keywords=="Campaign + Suicide") %>% #choose only one of the repeated datasets (tot long the same every time)
  summarize(milliontweets = round(sum(as.numeric(tot_n)/1000000), 2),
            meantweetsperday = round(mean(as.numeric(tot_n)), 2)) %>% 
  ungroup()
```
## Number of tweets per query per time period in Oregon

* Baseline: 24 September 2018 to 24 March 2019 (half a year ending 2 weeks before the campaign)
* Campaign-week: 7-14 April 2019
* After: 15 April until 15 October 2019 (half a year after the campaign)

```{r}
df %>% 
  filter(state=="Oregon") %>% 
  group_by(keywords, period) %>%
  summarize(meantweetsperday = round(mean(as.numeric(keywords_n)), 2),
            sd=round(sd(keywords_n), 2)) %>% 
  ungroup()
```

## Number of tweets per query per time period in Washington

```{r}
df %>% 
  filter(state=="Washington") %>% 
  group_by(keywords, period) %>%
  summarize(meantweetsperday = round(mean(as.numeric(keywords_n)), 2),
            sd=round(sd(keywords_n), 2)) %>% 
  ungroup()
```
